---
title: "R_project: Programmation et traitement statistique des données"
author: "MC-MF"
date: "2025-11-11"
output: html_document
---

# Question 1

$$\hat{f}^{',1}_n(x) = \frac{\hat{h}'_n(x)}{\hat{g}_n(x)} -  \frac{\hat{h}_n(x)\hat{g}'_n(x)}{\hat{g}_n(x)^2}$$
$$\hat{f}^{',2}_n(x) = \frac{\hat{h}'_n(x)}{g(x)} - \frac{\hat{h}_n(x)g'(x)}{g(x)^2}$$
$$\hat{f}^{',3}_n(x) = \frac{1}{n}\sum_{k=1}^n \frac{Y_k}{g(X_k)h_k^2}K'(\frac{x-X_k}{h_k})$$

$$\hat{h}_n(x) = \frac{1}{n} \sum_{k=1}^n \frac{Y_k}{h_k}K(\frac{x-X_k}{h_k})$$

$$\hat{h}'_n(x) = \frac{1}{n} \sum_{k=1}^n \frac{Y_k}{h_k^2}K'(\frac{x-X_k}{h_k})$$


$$\hat{g}_n(x) = \frac{1}{n}\sum_{k=1}^n\frac{1}{h_k}K(\frac{x-X_k}{h_k})$$

```{r}
N <- 10000 # sample size
alpha <- 0.3 # value given in the paper obtained with CV
h <- (1:N)^(-alpha)# bandwidth

h_N <- function(x) {
  return((1/N)*(sum((Y_N/h)*K((x-X_N)/h))))
}
g_N <- function(x) {
  return((1/N)*(sum((1/h)*K((x-X_N)/h))))
}
h_N_prime <- function(x) {
  return((1/N)*(sum((Y_N/h^2)*K_prime((x-X_N)/h))))
}
g_N_prime <- function(x) {
  return((1/N)*(sum((1/h^2)*K_prime((x-X_N)/h))))
}
es_n1 <- function(x) { # recursive version of Nadaraya Watson (Parzen 1962; Rosenblatt 1956)
  return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
}
es_n2 <- function(x) { # Johnston (1982)
  return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
}
es_n3 <- function(x) { # Wand and Jones (1995)
  return((1/N)*(sum((Y_N/(g(x)*h^2))*K_prime((x-X_N)/h))))
}
```

# Question 2: Simulations

## Bandwidth Selection with Cross-Validation

Alpha would depend on the chosen kernel (among gaussian, epanechnikov, quadratic) and on the chosen non parametric estimator.

Here we take K from Epanechnikov and we evaluate alpha depending on the chosen estimator in recursive form (original, Johnston, Wand)).

All along we will deal with $$X \sim U[0, 1] \Rightarrow \hat{f}^{',2} = \hat{f}^{',3}$$

```{r}
f <- function(x) { # primitive of target function
  return((x+2)*sin(4*pi*x^2)+2*sin(8*pi*x))
}
f_prime <- function(x) { # target function
  return(sin(4*pi*x^2)+8*pi*x*(x+2)*cos(4*pi*x^2) + 8*pi*cos(8*pi*x))
}
g <- function(x) { # density of X
  return(1)
}
g_prime <- function(x){
  return(0)
}
```

```{r}
# data generation
N <- 2500 # sample size
eps <- rnorm(N, 0, 1) # error term
X <- runif(N, 0, 1) # X ~ U[0, 1] with density g
Y <- f(X) + eps # (Y, X) non parametric relation
```

```{r}
K <- function(x) { # Epanechnikov kernel
  return(ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)) 
}
K_prime <- function(x) { # derivative of kernel
  return(ifelse(abs(x) <= 1, -(3/2)*x, 0))
}
```

```{r}
# Cross Validation function
CV_es1 <- function(alpha) { # CV for each estimator given K, g, f, X, Y
    cv_score <- 0
    h <- (1:N)^(-alpha)
    H <- h # saving for LOOCV
    for (k in 1:length(X)) { # proceeding to LOOCV
      # cv_[1](X_(-1)) + ...cv_[k](X_(-k) + ... + cv_[n](X_(-n))
      X_N <- X # (X_1, ..., X_N)
      Y_N <- Y # (Y_1, ..., Y_N)
      h <- H[-k] # leaving one out
      X_N <- X_N[-k] # leaving one out
      Y_N <- Y_N[-k] # leaving one out
      es_n1 <- function(x) {
        return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
      }
      h_N <- function(x) {# p(Y_N,X_N,h) => impacted by LOOCV
        return((1/(length(X_N)))*(sum((Y_N/h)*K((x-X_N)/h))))
      }
      g_N <- function(x) {# p(X_N,h) => impacted by LOOCV
        return((1/(length(X_N)))*(sum((1/h)*K((x-X_N)/h))))
      }
      h_N_prime <- function(x) {# p(Y_N,X_N,h) => impacted by LOOCV
        return((1/(length(X_N)))*(sum((Y_N/h^2)*K_prime((x-X_N)/h))))
      }
      g_N_prime <- function(x) {# p(X_N,h) => impacted by LOOCV
        return((1/(length(X_N)))*(sum((1/h^2)*K_prime((x-X_N)/h))))
      }
      cv_score <- cv_score + (1/N)*(es_n1(X[k]) - f_prime(X[k]))^2
    }
      return(cv_score)
  }

CV_es2 <- function(alpha) { # CV for each estimator given K, g, f, X, Y
    cv_score <- 0
    h <- (1:N)^(-alpha)
    H <- h # saving for LOOCV
    for (k in 1:length(X)) { # proceeding to LOOCV
      X_N <- X # (X_1, ..., X_N)
      Y_N <- Y # (Y_1, ..., Y_N)
      h <- H[-k] # leaving one out
      X_N <- X_N[-k] # leaving one out
      Y_N <- Y_N[-k] # leaving one out
      es_n2 <- function(x) {
        return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
      }
      h_N <- function(x) {# p(Y_N,X_N,h) => impacted by LOOCV
        return((1/length(X_N))*(sum((Y_N/h)*K((x-X_N)/h))))
      }
      g_N <- function(x) {# p(X_N,h) => impacted by LOOCV
        return((1/length(X_N))*(sum((1/h)*K((x-X_N)/h))))
      }
      h_N_prime <- function(x) {# p(Y_N,X_N,h) => impacted by LOOCV
        return((1/length(X_N))*(sum((Y_N/h^2)*K_prime((x-X_N)/h))))
      }
      g_N_prime <- function(x) {# p(X_N,h) => impacted by LOOCV
        return((1/length(X_N))*(sum((1/h^2)*K_prime((x-X_N)/h))))
      }
      cv_score <- cv_score + (1/N)*(es_n2(X[k]) - f_prime(X[k]))^2
    }
    return(cv_score)
  }
```

```{r}
# Optimal bandwidth for Kernel 1 and es_n1, es_n2, es_n3
library(stats)
res_es1 <- optim(par=0.3, fn = CV_es1, lower = 1/5, upper=1/2, method="L-BFGS-B")
alpha_star_es1 <-res_es1$par
alpha_star_es1
res_es2 <- optim(par=0.3, fn = CV_es2, lower = 1/5, upper=1/2, method="L-BFGS-B")
alpha_star_es2 <-res_es2$par
alpha_star_es2
```
```{r}
# global functions reinit
f <- function(x) {
  return((x+2)*sin(4*pi*x^2)+2*sin(8*pi*x))
}
f_prime <- function(x) {
  return(sin(4*pi*x^2)+8*pi*x*(x+2)*cos(4*pi*x^2) + 8*pi*cos(8*pi*x))
}

h_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
g_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
h_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
g_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
# estimators
es_n1 <- function(x) {
  return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
}
es_n2 <- function(x) {
  return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
}
# density of X
g <- function(x) {
  return(1)
}
g_prime <- function(x){
  return(0)
}
```


## Figure 1 (almost sure convergence of three estimators) with Epanechnikov Kernel

```{r}
library(gridExtra)
alpha <- alpha_star_es1
h <- (1:N)^(-alpha) # bandwidth
u1 <- es_n1(x)
alpha <- alpha_star_es2
u2 <- es_n2(x)
h <- (1:N)^(-alpha) # bandwidth
library(ggplot2)
x <- seq(0, 1, length.out = 1000) # plot on interval 0, 1
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), u1, u2),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
g1 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(EK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

alpha <-  1/3 # obtained by trial-error from alpha*L00CV from histograms plot
h <- (1:N)^(-alpha) # bandwidth
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), es_n1(x), es_n2(x)),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
g2 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(EK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

g3 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(EK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

grid.arrange(g1, g2, g3, ncol = 2)
```

```{r}
K <- function(x) { # gaussian kernel
  return(dnorm(x)) 
}
K_prime <- function(x) { # derivative
  return(-(x*dnorm(x)))
}
```

```{r}
# Optimal bandwidth for Kernel 2 and es_n1, es_n2, es_n3
# Optimal bandwidth for Kernel 1 and es_n1, es_n2, es_n3
library(stats)
res_es1 <- optim(par=0.3, fn = CV_es1, lower = 1/5, upper=1/2, method="L-BFGS-B")
alpha_star_es1 <-res_es1$par
alpha_star_es1
res_es2 <- optim(par=0.3, fn = CV_es2, lower = 1/5, upper=1/2, method="L-BFGS-B")
alpha_star_es2 <-res_es2$par
alpha_star_es2
```


## Figure 1 (almost sure convergence of three estimators) with gaussian kernel

```{r}
library(gridExtra)
alpha <- 0.5
h <- (1:N)^(-alpha) # bandwidth
u1 <- es_n1(x)
alpha <- 0.5
u2 <- es_n2(x)
h <- (1:N)^(-alpha) # bandwidth
library(ggplot2)
x <- seq(0, 1, length.out = 1000) # plot on interval 0, 1
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), u1, u2),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
g1 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(GK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

alpha <-  1/3 # obtained by trial-error from alpha*L00CV from histograms plot
h <- (1:N)^(-alpha) # bandwidth
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), es_n1(x), es_n2(x)),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
g2 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(GK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

g3 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(GK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

grid.arrange(g1, g2, g3, ncol = 2)
```


```{r}
K <- function(x) { # quadratic kernel
  return(ifelse(abs(x) <= 1, (15/16)*(1-x^2)^2, 0))
}
K_prime <- function(x) { # its derivative
  return(ifelse(abs(x) <= 1, -(15/4)*(x)*(1-x^2), 0))
}
```

```{r}
library(stats)
res_es1 <- optim(par=0.3, fn = CV_es1, lower = 1/5, upper=1/2, method="L-BFGS-B")
alpha_star_es1 <-res_es1$par
alpha_star_es1
res_es2 <- optim(par=0.3, fn = CV_es2, lower = 1/5, upper=1/2, method="L-BFGS-B")
alpha_star_es2 <-res_es2$par
alpha_star_es2
```

```{r}
library(gridExtra)
library(ggplot2)
alpha <- 0.418
h <- (1:N)^(-alpha) # bandwidth
u1 <- es_n1(x)
alpha <- 0.372
u2 <- es_n2(x)
h <- (1:N)^(-alpha) # bandwidth
x <- seq(0, 1, length.out = 1000) # plot on interval 0, 1
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), u1, u2),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
g1 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(CK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

alpha <-  1/3 # obtained by trial-error from alpha*L00CV from histograms plot
h <- (1:N)^(-alpha) # bandwidth
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), es_n1(x), es_n2(x)),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
g2 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(QK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

alpha <-  0.8 # obtained by trial-error from alpha*L00CV from histograms plot
h <- (1:N)^(-alpha) # bandwidth
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), es_n1(x), es_n2(x)),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
g3 <- ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(QK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()

grid.arrange(g1, g2, g3, ncol = 2)
```


## Simulation Implementation

```{r}
# APPLY CONSTRAINT ON ALPHA FOR THEOREM 3.2
# APPLY LOOCV with ES_N1, ES_N2, ES_N3, EK, GK, QK => alpha* = 1/3
N <- 20000 # sample size
alpha <- 1/3 # optimal bandwidth by LOOCV under theoretical constraints

h <- (1:N)^(-alpha) # bandwidth
```

```{r}
# data generation
eps_N <- rnorm(N, 0, 1) # error term
X_N <- runif(N, 0, 1) # X ~ U[0, 1]
Y_N <- f(X_N) + eps_N # Non parametric relation between X and Y
```

```{r}
# global functions and data generation
f <- function(x) {
  return((x+2)*sin(4*pi*x^2)+2*sin(8*pi*x))
}
f_prime <- function(x) {
  return(sin(4*pi*x^2)+8*pi*x*(x+2)*cos(4*pi*x^2) + 8*pi*cos(8*pi*x))
}

h_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
g_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
h_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
g_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
# estimators
es_n1 <- function(x) {
  return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
}
es_n2 <- function(x) {
  return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
}
# density of X
g <- function(x) {
  return(1)
}
g_prime <- function(x){
  return(0)
}
```

```{r}
K <- function(x) {# epanechnikov kernel
  return(ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)) 
}
K_prime <- function(x) { # derivative
  return(ifelse(abs(x) <= 1, -(3/2)*x, 0))
}
```

## Figure 1 (almost sure convergence of three estimators) with Epanechnikov Kernel

```{r}
alpha <-  1/3 # obtained by trial-error from alpha*L00CV from histograms plot
h <- (1:N)^(-alpha) # bandwidth
library(ggplot2)
x <- seq(0, 1, length.out = 1000) # plot on interval 0, 1
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), es_n1(x), es_n2(x)),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(EK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()
```

```{r}
alpha <- 1/3 # optimal bandwidth by LOOCV
h <- (1:N)^(-alpha) # bandwidth

K <- function(x) { # gaussian kernel
  return(dnorm(x)) 
}
K_prime <- function(x) { # derivative
  return(-(x*dnorm(x)))
}
```

## Figure 1 (almost sure convergence of three estimators) with Gaussian Kernel

```{r}
library(ggplot2)
x <- seq(0, 1, length.out = 1000) # plot on interval 0, 1
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), es_n1(x), es_n2(x)),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(GK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()
```

```{r}
K <- function(x) { # quadratic kernel
  return(ifelse(abs(x) <= 1, (15/16)*(1-x^2)^2, 0))
}
K_prime <- function(x) { # its derivative
  return(ifelse(abs(x) <= 1, -(15/4)*(x)*(1-x^2), 0))
}
```

## Figure 1 (almost sure convergence of three estimators) with Quadratic Kernel

```{r}
library(ggplot2)
x <- seq(0, 1, length.out = 1000) # plot on interval 0, 1
df <- data.frame( # plot of f', es_n1, es_n2, es_n3
  x = rep(x, 3),
  y = c(f_prime(x), es_n1(x), es_n2(x)),
  curve_name = factor(rep(c("Target","Estimator 1", "Estimator 2"), each = length(x)))
)
ggplot(df, aes(x = x, y = y, color = curve_name, linetype = curve_name)) + # specifying the ggplot ; curve distinguishable by color and linetype
  geom_line(size = 0.5) +
  geom_point( # adding points gives better look
    data = subset(df, curve_name != "Target"),
    size = 0.5,
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted")) + # setting of linetypes by curve
  labs(
    title = "(QK) Almost sure Convergence of fn'_1, fn'_2 to f'",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function" # obtain a single legend applying for both colors and linetype
  ) +
  theme_classic()
```

```{r}
# global functions and data generation
f <- function(x) {
  return((x+2)*sin(4*pi*x^2)+2*sin(8*pi*x))
}
f_prime <- function(x) {
  return(sin(4*pi*x^2)+8*pi*x*(x+2)*cos(4*pi*x^2) + 8*pi*cos(8*pi*x))
}
# density of X
g <- function(x) {
  return(1)
}
g_prime <- function(x){
  return(0)
}
```

```{r}
# Select Kernel for CLT result
# epanechnikov kernel
K <- function(x) {
  return(ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)) 
}
K_prime <- function(x) {
  return(ifelse(abs(x) <= 1, -(3/2)*x, 0))
}
```

## Figure 2 (CLT) with Epanechnikov Kernel

```{r}
N <- 2000
B <- 500
x0 <- 0.5
alpha <- 0.8
x1 <- 0.8 
par(mfrow=c(2,2))
set.seed(123)
ES1_X0 <- rep(NA, B)
ES2_X0 <- rep(NA, B)
ES1_X1 <- rep(NA, B)
ES2_X1 <- rep(NA, B)
for (b in 1:B) {
  h <- (1:N)^(-alpha)
  eps_N <- rnorm(N, 0, 1)
  X_N <- runif(N, 0, 1)
  Y_N <- f(X_N) + eps_N
  h_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  h_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  es_n1 <- function(x) {
    return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
    }
  es_n2 <- function(x) {
    return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
    }
  ES1_X0[b] <- sqrt(N*(1/N^alpha)^3)*
(es_n1(x0) - f_prime(x0))
  ES2_X0[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x0) - f_prime(x0))
  ES1_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n1(x1) - f_prime(x1))
  ES2_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x1) - f_prime(x1))
}

VES1X0 <- c(mean(ES1_X0),sd(ES1_X0))
VES1X1 <- c(mean(ES1_X1),sd(ES1_X1))
VES2X0 <- c(mean(ES2_X0),sd(ES2_X0))
VES2X1 <- c(mean(ES2_X1),sd(ES2_X1))
nbins <- 10
# estimator 1
hist(main="", ES1_X0, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES1_X0)$y)*1.5))
XVALS_ES1_X0 <- seq(min(ES1_X0), max(ES1_X0), length.out = 200)
lines(XVALS_ES1_X0, dnorm(XVALS_ES1_X0, VES1X0[1], VES1X0[2]), col = "red", lwd = 2)
hist(main="", ES1_X1, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES1_X1)$y)*1.5))
XVALS_ES1_X1 <- seq(min(ES1_X1), max(ES1_X1), length.out = 200)
lines(XVALS_ES1_X1, dnorm(XVALS_ES1_X1, VES1X1[1], VES1X1[2]), col = "red", lwd = 2)
# estimator 2
hist(main="", ES2_X0, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES2_X0)$y)*1.5))
XVALS_ES2_X0 <- seq(min(ES2_X0), max(ES2_X0), length.out = 200)
lines(XVALS_ES2_X0, dnorm(XVALS_ES2_X0, VES2X0[1], VES2X0[2]), col = "red", lwd = 2)
hist(main="", ES2_X1, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES2_X1)$y)*1.5))
XVALS_ES2_X1 <- seq(min(ES2_X1), max(ES2_X1), length.out = 200)
lines(XVALS_ES2_X1, dnorm(XVALS_ES2_X1, VES2X1[1], VES2X1[2]), col = "red", lwd = 2)
```

```{r}
K <- function(x) {# gaussian kernel
  return(dnorm(x)) 
}
K_prime <- function(x) {# deriv
  return(-(x*(K(x))))
}
```

## Figure 2 (CLT) with Gaussian Kernel

```{r}
N <- 2000
B <- 500
x0 <- 0.5
alpha <- 0.9
x1 <- 0.8
par(mfrow=c(3,2))

set.seed(123)
ES1_X0 <- rep(NA, B)
ES2_X0 <- rep(NA, B)
ES1_X1 <- rep(NA, B)
ES2_X1 <- rep(NA, B)
for (b in 1:B) {
  h <- (1:N)^(-alpha)
  eps_N <- rnorm(N, 0, 1)
  X_N <- runif(N, 0, 1)
  Y_N <- f(X_N) + eps_N
  h_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  h_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  es_n1 <- function(x) {
    return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
    }
  es_n2 <- function(x) {
    return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
    }
  ES1_X0[b] <- sqrt(N*(1/N^alpha)^3)*
(es_n1(x0) - f_prime(x0))
  ES2_X0[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x0) - f_prime(x0))
  ES1_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n1(x1) - f_prime(x1))
  ES2_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x1) - f_prime(x1))
}

VES1X0 <- c(mean(ES1_X0),sd(ES1_X0))
VES1X1 <- c(mean(ES1_X1),sd(ES1_X1))
VES2X0 <- c(mean(ES2_X0),sd(ES2_X0))
VES2X1 <- c(mean(ES2_X1),sd(ES2_X1))
nbins <- 10
XVALS_ES1_X0 <- seq(min(ES1_X0), max(ES1_X0), length.out = 200)
hist(main="", ES1_X0, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES1_X0)$y)*1.5))
lines(XVALS_ES1_X0, dnorm(XVALS_ES1_X0, VES1X0[1], VES1X0[2]), col = "red", lwd = 2)
hist(main="", ES1_X1, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES1_X1)$y)*1.5))
XVALS_ES1_X1 <- seq(min(ES1_X1), max(ES1_X1), length.out = 200)
lines(XVALS_ES1_X1, dnorm(XVALS_ES1_X1, VES1X1[1], VES1X1[2]), col = "red", lwd = 2)

XVALS_ES2_X0 <- seq(min(ES2_X0), max(ES2_X0), length.out = 200)
hist(main="", ES2_X0, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES2_X0)$y)*1.5))
lines(XVALS_ES2_X0, dnorm(XVALS_ES2_X0, VES2X0[1], VES2X0[2]), col = "red", lwd = 2)
hist(main="", ES2_X1, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES2_X1)$y)*1.5))
XVALS_ES2_X1 <- seq(min(ES2_X1), max(ES2_X1), length.out = 200)
lines(XVALS_ES2_X1, dnorm(XVALS_ES2_X1, VES2X1[1], VES2X1[2]), col = "red", lwd = 2)
```

```{r}
# quadratic kernel
K <- function(x) {
  return(ifelse(abs(x) <= 1, (15/16)*(1-x^2)^2, 0))
}
K_prime <- function(x) {
  return(ifelse(abs(x) <= 1, -(15/4)*(x)*(1-x^2), 0))
}
```

## Figure 2 (CLT) with Quadratic Kernel

```{r}
N <- 2000
B <- 500
x0 <- 0.5
alpha <- 0.83 # to obtain centered for both x0 and x1; though not a.s. conv on x1 with alpha != 1/3 high variance may explain why centered ? x0 centered on a.s. conv plot at alpha in (0.319 +- epsilon) and f^2(x0) = 0, see normality convergence on paper
x1 <- 0.8
par(mfrow=c(3,2))

set.seed(123)
ES1_X0 <- rep(NA, B)
ES2_X0 <- rep(NA, B)
ES1_X1 <- rep(NA, B)
ES2_X1 <- rep(NA, B)
for (b in 1:B) {
  h <- (1:N)^(-alpha)
  eps_N <- rnorm(N, 0, 1)
  X_N <- runif(N, 0, 1)
  Y_N <- f(X_N) + eps_N
  h_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  h_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  es_n1 <- function(x) {
    return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
    }
  es_n2 <- function(x) {
    return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
    }
  ES1_X0[b] <- sqrt(N*(1/N^alpha)^3)*
(es_n1(x0) - f_prime(x0))
  ES2_X0[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x0) - f_prime(x0))
  ES1_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n1(x1) - f_prime(x1))
  ES2_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x1) - f_prime(x1))
}

VES1X0 <- c(mean(ES1_X0),sd(ES1_X0))
VES1X1 <- c(mean(ES1_X1),sd(ES1_X1))
VES2X0 <- c(mean(ES2_X0),sd(ES2_X0))
VES2X1 <- c(mean(ES2_X1),sd(ES2_X1))
nbins <- 10
XVALS_ES1_X0 <- seq(min(ES1_X0), max(ES1_X0), length.out = 200)
hist(main="", ES1_X0, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES1_X0)$y)*1.5))
lines(XVALS_ES1_X0, dnorm(XVALS_ES1_X0, VES1X0[1], VES1X0[2]), col = "red", lwd = 2)
hist(main="", ES1_X1, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES1_X1)$y)*1.5))
XVALS_ES1_X1 <- seq(min(ES1_X1), max(ES1_X1), length.out = 200)
lines(XVALS_ES1_X1, dnorm(XVALS_ES1_X1, VES1X1[1], VES1X1[2]), col = "red", lwd = 2)

XVALS_ES2_X0 <- seq(min(ES2_X0), max(ES2_X0), length.out = 200)
hist(main="", ES2_X0, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES2_X0)$y)*1.5))
lines(XVALS_ES2_X0, dnorm(XVALS_ES2_X0, VES2X0[1], VES2X0[2]), col = "red", lwd = 2)
hist(main="", ES2_X1, breaks=nbins, freq = FALSE, col = "blue",
     xlab = "", ylim = c(0, max(density(ES2_X1)$y)*1.5))
XVALS_ES2_X1 <- seq(min(ES2_X1), max(ES2_X1), length.out = 200)
lines(XVALS_ES2_X1, dnorm(XVALS_ES2_X1, VES2X1[1], VES2X1[2]), col = "red", lwd = 2)

```
## Question 3: Comparing Kernels with MSE criterion
## x = 0.5 and x = 0.8
## Question : Comparing Kernels with MISE criterion

```{r}
set.seed(123)
# Setting

# data generation parameters
alpha <- 1/3 # optimal bandwidth by LOOCV
h <- (1:N)^(-alpha) # bandwidth
N <- 1000 # sample size per batch
B <- 500 # batch

x = c(0.5, 0.8) # values for MSE
d <- rep(1/N, N) # discretizing tool for MISE
x_d <- (1:N)/N # x_val for MISE
```

```{r}
# global functions and data generation
f <- function(x) {
  return((x+2)*sin(4*pi*x^2)+2*sin(8*pi*x))
}
f_prime <- function(x) {
  return(sin(4*pi*x^2)+8*pi*x*(x+2)*cos(4*pi*x^2) + 8*pi*cos(8*pi*x))
}

h_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
g_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
h_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
g_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
# estimators
es_n1 <- function(x) {
  return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
}
es_n2 <- function(x) {
  return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
}
# density of X
g <- function(x) {
  return(1)
}
g_prime <- function(x){
  return(0)
}
```


```{r}
# Kernel 1 EK
K <- function(x) {
  return(ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)) 
}
K_prime <- function(x) {
  return(ifelse(abs(x) <= 1, -(3/2)*x, 0))
}

ES1_X0_K1 <- rep(NA, B)
ES2_X0_K1 <- rep(NA, B)
ES1_X1_K1 <- rep(NA, B)
ES2_X1_K1 <- rep(NA, B)
for (b in 1:B) {
  eps_N <- rnorm(N, 0, 1) # error term
  X_N <- runif(N, 0, 1) # X ~ U[0, 1]
  Y_N <- f(X_N) + eps_N # Non parametric relation between X and Y
  ES1_X0_K1[b] <- es_n1(x0) 
  ES2_X0_K1[b] <- es_n2(x0)
  ES1_X1_K1[b] <- es_n1(x1)
  ES2_X1_K1[b] <- es_n2(x1)
}
MSE_ES1_K1_X0 = mean((ES1_X0_K1 - f_prime(x0))^2) + var(ES1_X0_K1)
MSE_ES2_K1_X0 = mean((ES2_X0_K1 - f_prime(x0))^2) + var(ES2_X0_K1)
MSE_ES1_K1_X1 = mean((ES1_X1_K1 - f_prime(x1))^2) + var(ES1_X1_K1)
MSE_ES2_K1_X1 = mean((ES2_X1_K1 - f_prime(x1))^2) + var(ES2_X1_K1)

# Kernel 2 GK

K <- function(x) {
  return(dnorm(x)) 
}
K_prime <- function(x) {
  return(-(x*(K(x))))
}

alpha <- 1/3

ES1_X0_K2 <- rep(NA, B)
ES2_X0_K2 <- rep(NA, B)
ES1_X1_K2 <- rep(NA, B)
ES2_X1_K2 <- rep(NA, B)
for (b in 1:B) {
  e <- rnorm(N, 0, 1)
  X <- runif(N, 0, 1)
  Y <- f(X) + e
  ES1_X0_K2[b] <- es_n1(x0)
  ES2_X0_K2[b] <- es_n2(x0)
  ES1_X1_K2[b] <- es_n1(x1)
  ES2_X1_K2[b] <- es_n2(x1)
}
MSE_ES1_K2_X0 = mean((ES1_X0_K2 - f_prime(x0))^2) + var(ES1_X0_K2)
MSE_ES2_K2_X0 = mean((ES2_X0_K2 - f_prime(x0))^2) + var(ES2_X0_K2)
MSE_ES1_K2_X1 = mean((ES1_X1_K2 - f_prime(x1))^2) + var(ES1_X1_K2)
MSE_ES2_K2_X1 = mean((ES2_X1_K2 - f_prime(x1))^2) + var(ES2_X1_K2)

# Kernel 3 QK

K <- function(x) {
  return(ifelse(abs(x) <= 1, (15/16)*(1-x^2)^2, 0))
}
K_prime <- function(x) {
  return(ifelse(abs(x) <= 1, -(15/4)*(x)*(1-x^2), 0))
}

ES1_X0_K3 <- rep(NA, B)
ES2_X0_K3 <- rep(NA, B)
ES1_X1_K3 <- rep(NA, B)
ES2_X1_K3 <- rep(NA, B)
for (b in 1:B) {
  e <- rnorm(N, 0, 1)
  X <- runif(N, 0, 1)
  Y <- f(X) + e
  ES1_X0_K3[b] <- es_n1(x0)
  ES2_X0_K3[b] <- es_n2(x0)
  ES1_X1_K3[b] <- es_n1(x1)
  ES2_X1_K3[b] <- es_n2(x1)
}
MSE_ES1_K3_X0 = mean((ES1_X0_K3 - f_prime(x0))^2) + var(ES1_X0_K3)
MSE_ES2_K3_X0 = mean((ES2_X0_K3 - f_prime(x0))^2) + var(ES2_X0_K3)
MSE_ES1_K3_X1 = mean((ES1_X1_K3 - f_prime(x1))^2) + var(ES1_X1_K3)
MSE_ES2_K3_X1 = mean((ES2_X1_K3 - f_prime(x1))^2) + var(ES2_X1_K3)

print("MSE for K1")
c(MSE_ES1_K1_X0, MSE_ES2_K1_X0, MSE_ES1_K1_X1, MSE_ES2_K1_X1)
print("MSE for K2")
c(MSE_ES1_K2_X0, MSE_ES2_K2_X0, MSE_ES1_K2_X1, MSE_ES2_K2_X1)
print("MSE for K3")
c(MSE_ES1_K3_X0, MSE_ES2_K3_X0, MSE_ES1_K3_X1, MSE_ES2_K3_X1)
# Interpretation for x0: Optimality for x0 with epanechnikov kernel and Estimator 1 with alpha btw [1/5, 1/3]
# Interpretation for x1: Optimality for x1 with quadratic kernel and Estimator 1 with alpha btw [1/5, 1/3];
```
